---
title: "Setting Up R Assignment"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

```{r}
library("ISLR")
```
#ISLR package has been installed using install.packages("ISLR")
#calling the library using command library("ISLR") and activating the installed ISLR package

```{r}
summary(Carseats)
```
#Printing out the summary of carseats dataset

```{r}
nrow(Carseats)
```
#printing out the count of rows that are present in the carseats dataset overall
```{r}
max(Carseats$Advertising)
```
#determining the max value in the advertising attribute in carseats dataset
#
```{r}
IQR(Carseats$Price)
```
#printing out the Interquartile Range of the price attribute



```{r}
library(ggplot2)
ggplot(Carseats)+
  aes(
    x = Price,
    y = Sales
  )+
  geom_point(shape="square", size=1.4)+theme_minimal()
```

#Plotting the sales against the price using ggplot

#It can be found from the plot that
#x and y that is price and sales variables of carseats dataset have negative or inverse relationship
#There is a linear relationship between two variables x and y 
#It can also be seen that the points are scattered and are not too close which indicates that the relationship between the variables is not too strong.
```{r}
cor(Carseats$Price, Carseats$Sales)
```
#Printing out the correlation of two attributes price and sales
#The calculated correlation value of two attributes price and sales is -0.4449507 which suggests that the two attributes are negatively correlated.

